---
title: "GaussQuad"
author: "Yomna Eid"
date: "2025-01-15"
output: html_document
---

Import dataset 1:
```{r}
library(stars)

r <- read_stars(r"(C:\Users\yeid\Documents\Dead Laptop Files\Documents\1b. PhD\2. On-going Publications\1. Downsampling\Downsampled_Output\2006\impervious2006_20m.tif)", na.rm = TRUE) |> setNames("IMP2006") %>% st_as_stars() %>% na.omit()

p = st_as_sfc(st_bbox(r))
```

Compute the variance of the raster image:
```{r}
(v = var(as.vector(r[[1]]), na.rm = TRUE))
```

The different grid sizes for downsampling:
```{r}
gridsizes = c(20, 40, 80, 100, 200, 400, 800, 1000, 2000, 4000, 8000, 10000)
thinning = gridsizes / 10
```

Compute the standard errors associated with the different grid sizes, for
random sampling:
```{r}
(n = prod(dim(r)) / (thinning^2))
(se.random_sampling = sqrt(v/n))
names(se.random_sampling) = gridsizes
se.random_sampling
```

Now we fit the variogram to our raster image, to identify the model for the covariance function:
```{r}
library(gstat)
set.seed(1353)
pts = st_sample(p, 35000)
r.sample = st_extract(r, pts) %>% na.omit()
dist = gridsizes
v = variogram(IMP2006~1, r.sample, boundaries = dist)
plot(v)
m = "Exp"
v0 = vgm(.1, m, 100)
(v.model = fit.variogram(v, vgm(.1, m, 500, add.to = v0)))
plot(v, v.model)

v.model
```
Now we try two methodologies for computing the covariance function and the standard error using Ripley's 1981:

1. MONTE-CARLO INTEGRATION

Compute the mean point-point covariance function value, using random sampling (Monte-Carlo integration):

```{r}
n = 1e4
make_unif_pts = function(bbox, n) {
	x = runif(n, bbox["xmin"], bbox["xmax"])
	y = runif(n, bbox["ymin"], bbox["ymax"])
	as.matrix(cbind(x = x, y = y))
}

mean_cov = function(x, y, model) {
	d = if (missing(y))
			as.vector(as.matrix(dist(x)))
		else {
			all = rbind(x, y)
			n1 = nrow(x)
			n2 = nrow(y)
			as.vector(as.matrix(dist(all))[(n1+1):(n1+n2),1:n1])
		}
	mean(variogramLine(v.model, covariance = TRUE, dist_vector = d, n)$gamma)
}

x = make_unif_pts(st_bbox(r), n)
y = make_unif_pts(st_bbox(r), n)
(C_mean = mean_cov(x, y, v.model))
```

Create a function that computes the SE using Ripley's 1981 eq. 3.4 (page 23), and use it:

```{r}
make_grid = function(size, bbox, n = Inf) {
	x = seq(bbox["xmin"], bbox["xmax"], size)
	y = seq(bbox["ymin"], bbox["ymax"], size)
	grd = as.matrix(expand.grid(x = x, y = y))
	if (n < nrow(grd))
		grd = grd[sample(nrow(grd), n), ]
	grd
}

compute_se = function(size, model, maxpts = 1e4, C_mean) {
	grd = make_grid(size, st_bbox(r), maxpts)
	pts = make_unif_pts(st_bbox(r), maxpts)
	# eq 3.4, RHS part:
	var = mean_cov(grd, model = v.model) -
		2 * mean_cov(grd, pts, v.model) + C_mean
	sqrt(var)
}

#sel = -(1:5)
sel = TRUE
se = sapply(gridsizes[sel], compute_se, model = v.model, C_mean = C_mean)
names(se) = gridsizes[sel]
se
se.random_sampling 
```

2. GAUSS-QUADRATURE

Now compute the covariance function values, using deterministic (rather than stochastic - above) method of Gauss Quadrature, using systematic/symmetric sampling:

```{r}
# Define Gaussian quadrature points and weights

#n = 4
gauss <- c(-0.8611363116, -0.3399810436, 0.3399810436, 0.8611363116)
gauss_w <- c(0.3478548452, 0.6521451548, 0.6521451548, 0.3478548452)
 
#n = 6
gauss <- c(-0.9324695142, -0.6612093864, -0.2386191860, 0.2386191860, 0.6612093864, 0.9324695142)
gauss_w <- c(0.1713244924, 0.3607615730, 0.4679139346, 0.4679139346, 0.3607615730, 0.1713244924)

#n = 8
gauss <- c(-0.96028985649753623168, -0.79666647741362673959, -0.52553240991632898582, -0.18343464249564980494, 0.18343464249564980494, 0.52553240991632898582, 0.79666647741362673959, 0.96028985649753623168)
gauss_w <- c(0.10122853629037625915, 0.22238103445337447054, 0.31370664587788728734, 0.36268378337836198297, 0.36268378337836198297, 0.31370664587788728734, 0.22238103445337447054, 0.10122853629037625915)

#n = 10
gauss <- c(-0.97390652851717172008, -0.86506336668898451073, -0.67940956829902440623, -0.43339539412924719080, -0.14887433898163121089, 0.14887433898163121089, 0.43339539412924719080, 0.67940956829902440623, 0.86506336668898451073, 0.97390652851717172008)
gauss_w <- c(0.06667134430868813759, 0.14945134915058059315, 0.21908636251598204400, 0.26926671930999635509, 0.29552422471475287017, 0.29552422471475287017, 0.26926671930999635509, 0.21908636251598204400, 0.14945134915058059315, 0.06667134430868813759)

#n = 20
gauss <- c(-0.99312859918509492479, -0.96397192727791379127, -0.91223442825132590587, -0.83911697182221882339, -0.74633190646015079261, -0.63605368072651502545, -0.51086700195082709800, -0.37370608871541956067, -0.22778585114164507808, -0.07652652113349733375, 0.07652652113349733375, 0.22778585114164507808, 0.37370608871541956067, 0.51086700195082709800, 0.63605368072651502545, 0.74633190646015079261, 0.83911697182221882339, 0.91223442825132590587, 0.96397192727791379127, 0.99312859918509492479)
gauss_w <- c(0.01761400713915211831, 0.04060142980038694133, 0.06267204833410906357, 0.08327674157670474873, 0.10193011981724043504, 0.11819453196151841731, 0.13168863844917662690, 0.14209610931838205133, 0.14917298647260374679, 0.15275338713072585070, 0.15275338713072585070, 0.14917298647260374679, 0.14209610931838205133, 0.13168863844917662690, 0.11819453196151841731, 0.10193011981724043504, 0.08327674157670474873, 0.06267204833410906357, 0.04060142980038694133, 0.01761400713915211831)
```

First, we compute the Block-Sample Covariance using GQ-2D:

(Testing with a Pure Nugget model)
```{r}
# Define the covariance function for a cumulative nugget effect
cov_func <- function(x0, y0, x1, y1) {
  # Compute h as the Euclidean distance
  h <- sqrt((x1 - x0)**2 + (y1 - y0)**2)
  # Calculate the covariance: equals nugget_sill when h = 0, otherwise 0
  if (h == 0) {
    return(1) # Assuming this is the nugget effect variance
  } else {
    return(0)
  }}

# Function to perform Gaussian quadrature in 2D
gaussian_quadrature_2d <- function(cov_func, r) {

  # Extract integration limits
  bbox <- st_bbox(r)
  x_min <- bbox["xmin"]
  x_max <- bbox["xmax"]
  y_min <- bbox["ymin"]
  y_max <- bbox["ymax"]
  
  # Transform quadrature points to the integration domain
  transform_x <- function(xi) {
    0.5 * (x_max - x_min) * xi + 0.5 * (x_max + x_min)
  }
  
  transform_y <- function(yi) {
    0.5 * (y_max - y_min) * yi + 0.5 * (y_max + y_min)
  }
  
  transform_wx <- function(weight){
    0.5 * (x_max - x_min) * weight
  }

  transform_wy <- function(weight){
    0.5 * (y_max - y_min) * weight
  }

  # Transform gauss weights to the integration domain
  #gauss_wx <- 0.5 * (x_max - x_min) * gauss_w
  #gauss_wy <- 0.5 * (y_max - y_min) * gauss_w

  # Initialize integral value
  integral_value <- 0

  # Example Sample point
  x0 = 9
  y0 = 3

  # Perform double summation over Gaussian quadrature points
  for (i in 1:length(gauss)) {
    for (j in 1:length(gauss)) {
      # Map quadrature points to the integration domain
      x1 <- transform_x(gauss[i])  # First point in x
      y1 <- transform_y(gauss[j])  # First point in y
      gauss_wx <- transform_wx(gauss_w[i])
      gauss_wy <- transform_wy(gauss_w[j])
      # Evaluate the covariance function using h
      cov_val <- cov_func(x0, y0, x1, y1)
      integral_value <- integral_value +  gauss_wx * gauss_wy * cov_val
      # debug using print(paste("x0:", x0, "y0:", y0, "x1:", x1, "y1:", y1, "cov_val:", cov_val, "integral_value:", integral_value))
        }
      }
  # Scale by the area of the block
  area <- (x_max - x_min) * (y_max - y_min)
  integral_value <- integral_value / area
  return(integral_value)
}

# Compute the integral
result <- gaussian_quadrature_2d(cov_func, sample_data)

# Print the result
cat("The integral of the covariance function over the domain is:", result, "\n")
```

Next, we compute the Block-Block Covariance by extending the GQ-2D:

```{r}
# Define the covariance function for a cumulative nugget effect
cov_func <- function(x2, y2, x1, y1) {
  # Compute h as the Euclidean distance
  h <- sqrt((x1 - x2)**2 + (y1 - y2)**2)
  # # Testing by calculating the covariance value for a pure nugget model (cov equals nugget_sill when h = 0, otherwise 0)
  #   if (h < 1e-8) {
  #     return(0)
  #   } else {
  #     return(variogramLine(v.model, covariance = TRUE, dist_vector = h)$gamma)
  #     }
  # Double Exponential covariance function
  term1 <- v.model$psill[1] * (exp(-h/v.model$range[1]))
  term2 <- v.model$psill[2] * (exp(-h/v.model$range[2]))
  return(term1 + term2) 
  }

# Function to perform Gaussian quadrature in 2D
gaussian_quadrature_2d <- function(cov_func, r) {

  # Extract integration limits
  bbox <- st_bbox(r)
  x_min <- bbox["xmin"]
  x_max <- bbox["xmax"]
  y_min <- bbox["ymin"]
  y_max <- bbox["ymax"]

  # Transform quadrature points to the integration domain
  transform_x <- function(xi) {
    0.5 * (x_max - x_min) * xi + 0.5 * (x_max + x_min)
  }
  transform_y <- function(yi) {
    0.5 * (y_max - y_min) * yi + 0.5 * (y_max + y_min)
  }
  transform_wx <- function(weight){
    0.5 * (x_max - x_min) * weight
  }
  transform_wy <- function(weight){
    0.5 * (y_max - y_min) * weight
  }
  
  # Transform gauss weights to the integration domain
  #gauss_wx <- 0.5 * (x_max - x_min) * gauss_w
  #gauss_wy <- 0.5 * (y_max - y_min) * gauss_w

  # Initialize integral value
  integral_value <- 0

  # Perform double summation over Gaussian quadrature points
  for (i in seq_along(gauss)) {
    for (j in seq_along(gauss)) {
      # Map quadrature points to the integration domain
      x1 <- transform_x(gauss[i])  # First point in x
      y1 <- transform_y(gauss[j])  # First point in y
      gauss_wx1 <- transform_wx(gauss_w[i])
      gauss_wy1 <- transform_wy(gauss_w[j])
      
      first_integral <- 0

      for (k in seq_along(gauss)) {
        for (l in seq_along(gauss)) {
     	  # Map quadrature points to the integration domain
      	  x2 <- transform_x(gauss[k])  # First point in x
	  y2 <- transform_y(gauss[l])  # First point in y
      	  gauss_wx2 <- transform_wx(gauss_w[k])
      	  gauss_wy2 <- transform_wy(gauss_w[l])
  	  cov_val <- cov_func(x2, y2, x1, y1)

      	  first_integral <- first_integral + gauss_wx2 * gauss_wy2 * cov_val 
      
		#print(paste(gauss_wy2, "first_integral:", first_integral))
      		#print(paste("x2:", x2, "y2:", y2, "x1:", x1, "y1:", y1, "cov_val:", cov_val, "integral_value:", integral_value))
	  }
	}
      integral_value <- integral_value + first_integral * gauss_wx1 * gauss_wy1 
      }
    }
  # Scale by the area of the block
  area <- (x_max - x_min) * (y_max - y_min)
  integral_value <- integral_value / area**2
  return(integral_value)
}

# Compute the integral
result <- gaussian_quadrature_2d(cov_func, r)

# Print the result
cat("The integral of the covariance function over the domain is:", result, "\n")
```

Save:
```{r}
rm(r) # big
save.image("sampling.rda")
```

##########################TRASCH_CODE###################################

```{r}
dx = (bbox["xmax"] - bbox["xmin"])
dy = (bbox["ymax"] - bbox["ymin"])
x_mid = bbox["xmin"]+dx/2
y_mid = bbox["ymin"]+dy/2

n = 4
t = c(-0.4305681558,
		-0.1699905218,
		0.1699905218,
		0.4305681558)

w = c(0.1739274226,
		0.3260725774,
		0.3260725774,
		0.1739274226)

for (i in n){
  t_y = y_mid + t[i]*dy/2
  t_x = x_mid + t[i]*dx/2
  w_y = w[i]*dy/2
  w_x = w[i]*dx/2
}


d = if (missing(y))
			as.vector(as.matrix(dist(x))) else {
			all = rbind(x, y)
			n1 = nrow(x)
			n2 = nrow(y)
			as.vector(as.matrix(dist(all))[(n1+1):(n1+n2),1:n1])
		}

cov = v.model$psill[1] - v.model$psill[1]*(1-exp(-d/v.model$range[1])) + v.model$psill[2] - v.model$psill[2]*(1-exp(-d/v.model$range[2]))
cov

cov = 0.05682120 - 0.05682120*(1-exp(-abs(x-y)/81.30253)) + 0.04176899 - 0.04176899*(1-exp(-abs(x-y)/1182.16206))
```

```{r}
install.packages("statmod")
library(statmod)

gauss_legendre_2d_integration <- function(f, n_x, n_y) {
  quad_x <- gauss.quad(n_x, kind = "legendre")
  quad_y <- gauss.quad(n_y, kind = "legendre")
  integral_approximation <- 0
  
  for (i in 1:n_x) {
    for (j in 1:n_y) {
      integral_approximation <- integral_approximation +
        quad_x$weights[i] * quad_y$weights[j] * f(quad_x$nodes[i], quad_y$nodes[j])
    }
  }
  
  return(integral_approximation)
}

# Example function for integrating e^(abs(x-y))
f <- function(x, y) {
  return(exp(abs(x - y)))  
}

result <- gauss_legendre_2d_integration(f, n_x = 5, n_y = 5)
cat("The approximate double integral is:", result, "\n")
```

```{r}
# Number of nodes for each dimension
n <- 1e4

gauss_legendre <- function(n) {
  nodes <- rep(0, n)
  weights <- rep(0, n)
  
  for (i in 1:n) {
    xi <- cos(pi * (i - 0.25) / (n + 0.5))
    
    while(TRUE) {
      p1 <- 1
      p2 <- 0
      
      # Calculate Legendre polynomial values
      for (j in 1:n) {
        p3 <- p2
        p2 <- p1
        p1 <- ((2 * j - 1) * xi * p2 - (j - 1) * p3) / j
      }
      
      # Derivative of the polynomial
      pp <- n * (xi * p1 - p2) / (xi^2 - 1)
      
      # Newton's method for finding nodes
      xi1 <- xi
      xi <- xi1 - p1 / pp
      
      if (abs(xi - xi1) < 1e-10) break
    }
    
    nodes[i] <- xi
    weights[i] <- 2 / ((1 - xi^2) * pp^2)
  }
  
  return(list(nodes = nodes, weights = weights))
}

# Transform nodes and weights to any interval [a, b]
transform_gauss_legendre <- function(a, b, nodes, weights) {
  transformed_nodes <- (b - a) / 2 * nodes + (b + a) / 2
  transformed_weights <- (b - a) / 2 * weights
  return(list(nodes = transformed_nodes, weights = transformed_weights))
}

# Define your intervals [a, b] for x and [c, d] for y
bbox <- st_bbox(r)
a <- bbox["xmin"]
b <- bbox["xmax"]
c <- bbox["ymin"]
d <- bbox["ymax"]

# Get nodes and weights for x and y
gauss_result_x <- gauss_legendre(n)
transformed_x <- transform_gauss_legendre(a, b, gauss_result_x$nodes, gauss_result_x$weights)

gauss_result_y <- gauss_legendre(n)
transformed_y <- transform_gauss_legendre(c, d, gauss_result_y$nodes, gauss_result_y$weights)

nodes_x <- transformed_x$nodes
weights_x <- transformed_x$weights

nodes_y <- transformed_y$nodes
weights_y <- transformed_y$weights

m <- as.matrix(cbind(x = nodes_x, y = nodes_y))

mean_cov = function(x, y, model) {
	d = if (missing(y))
			as.vector(as.matrix(dist(x)))
		else {
			all = rbind(x, y)
			n1 = nrow(x)
			n2 = nrow(y)
			as.vector(as.matrix(dist(all))[(n1+1):(n1+n2),1:n1])
		}
	mean(variogramLine(v.model, covariance = TRUE, dist_vector = d, n)$gamma)
}

mean_cov = function(x, y, model) {
	d = if (missing(y))
			as.vector(as.matrix(dist(x)))
		else {
			all = rbind(x, y)
			n1 = nrow(x)
			n2 = nrow(y)
			as.vector(as.matrix(dist(all))[(n1+1):(n1+n2),1:n1])
		}
	mean(variogramLine(v.model, covariance = TRUE, dist_vector = d, n)$gamma)
}

f <- mean_cov
 
# Compute the double integral
approximation <- 0
for (i in 1:n) {
  for (j in 1:n) {
    approximation <- approximation + weights_x[i] * weights_y[j] * f(m[i], m[j],v.model)
  }
}

cat("Approximated double integral:", approximation, "\n")
```
